Privacy-Preserving Localization Using Private Linear-Combination Aggregation

### 2021-05-20 main.tex
Draft for estimation part of the problem formulation complete

@@

\section{Problem Statement} \label{sec:problem_statement}
In this work, we consider the context of privacy-presering range-sensor navigation, in which we wish for no sensor to learn information about the navigator or other sensors beyond their local measurements, and the navigator to learn no information about individual sensors beyond its location estimate. The problem is two-fold, in that we require explicit cryptographic requirements with a suitable encryption scheme meeting them, as well as an estimation scheme that can use the scheme in the context of range-sensor navigation.

@@

\subsection{Estimation problem} \label{subsec:est_problem}

The localization scenario we consider in this work is that of Bayesian self-navigation using range-only sensors. We consider localization in the two-dimensional case for simplicity but will derive methods suitable for an extension to the three-dimensional equivalent.

The navigator state captures the position and change in position ($\Delta$), and is of the form 
\begin{equation}
    \vec{x} = 
    \begin{bmatrix}
        x & \Delta x & y & \Delta y
    \end{bmatrix}^\top\,. \label{eqn:state_def}
\end{equation}
A known process model is followed, in which the state at time $k$ is given by
\begin{equation}
    \rvec{x}_k = \vec{f}(\rvec{x}_{k-1})+\rvec{w}_k\,, \label{eqn:process_model}
\end{equation}
with zero-mean Gaussian process noise $\rvec{w}_k \sim \mathcal{N}(\vec{0},\mat{Q})$. The measurement model is dependent on sensor $i$ and given by
\begin{equation}
    \rv{z}_{k,i} = h_i(\rvec{x}_k)+\rv{v}_{k,i}\,, \label{eqn:measurement_model}
\end{equation}
with noise $\rv{v}_{k,i} \sim \mathcal{N}(0,r_i)$. The measurement function $h_i$, for sensor $i$ at location
\begin{equation}
    \vec{s}_i = 
    \begin{bmatrix}
        s_{x,i} & s_{y,i}
    \end{bmatrix}^\top\,,
\end{equation} 
is defined as
\begin{equation}
    \begin{split}
        h_i(\vec{x}) &= \left\lVert
        \begin{bmatrix}
            x & y
        \end{bmatrix}^\top
        - \vec{s}_{i}\right\rVert \\
        &= \sqrt{(x-s_{x,i})^2 + (y-s_{y,i})^2}\,.
    \end{split}
\end{equation}
We wish to run a state estimation filter with models \eqref{eqn:process_model} and \eqref{eqn:measurement_model} such that all involved sensors $1 \leq i \leq n$ do not learn state estimates of $\rvec{x}_{k}$ and the navigator does not learn sensor locations $\vec{s}_i,\,1 \leq i \leq n$, measurement variances $r_i$ or their measurements $\rv{z}_{k,i}$ at any time $k$. We motivate these goals with the example of aircraft navigation in the presence of privately-operated range-measuring stations. It is reasonable to assume that the current state of an aircraft may not wish to be disclosed to unknown parties and, similarly, that tower locations may wish to be kept private from unidentified navigating aircraft. The additional safety to passengers provided by accurate aircraft localization, however, may be a goal of all those involved, justifying their cooperation.

% LEAKAGE

Leakage of information will be discussed in more detail in Section \ref{subsec:leakage}, however, we stress that in any estimation problem where process and measurement models are known, knowledge of sequential state estimates as is the case at the navigator, or the knowledge of one or more measurements as is the case by one or more malicious sensors, naturally leads to the leakage of some information. For this reason, we accept that average sensor information may be leaked to the navigator and that less accurate estimates of location may be leaked to sensors. We will instead show that individual sensor information remains private to sensors and that the most precise localization estimate, requiring all measurements, remains private to the navigator.

This has been captured in the associated secruity notions i nthe communication protocol

@@

Recalling the state definition \eqref{eqn:state_def} and two-dimensional measurement model \eqref{eqn:measurement_model}, we can see that $h_i$ \textit{cannot} be rearranged to a linear combination of elements private to the navigator. In addition, the Jacobian of $h_i$ at $\rvec{x}_{k|k-1}$,
\begin{equation}
    \mat{H}_{k,i} = 
    \begin{bmatrix}
        \frac{\rv{x}_{k|k-1} - s_{x,i}}{\sqrt{(\rv{x}_{k|k-1} - s_{x,i})^2 + (\rv{y}_{k|k-1} - s_{y,i})^2}} \\
        \frac{\rv{y}_{k|k-1} - s_{y,i}}{\sqrt{(\rv{x}_{k|k-1} - s_{x,i})^2 + (\rv{y}_{k|k-1} - s_{y,i})^2}} 
    \end{bmatrix}\,,
\end{equation}
cannot be either. Therefore, we consider the modified measurement functions
\begin{equation}
    h'_i(\vec{x}) = h_i(\vec{x})^2\,. \label{eqn:modified_measurement_func}
\end{equation}
A measurement function in this form allows rearrangement to a linear combination of navigator information, and thus of the corresponding modified measurement vector $\rvec{i}'_{k,i}$ and measurement matrix $\mat{I}'_{k,i}$ as well. The modified measurement functions $h'_i$ can be written as
\begin{equation}
    \begin{split}
        h'_i(\vec{x}) &= \left\lVert
        \begin{bmatrix}
            x & y
        \end{bmatrix}^\top - \vec{s}_i\right\rVert^2 \\
        &= (x - s_{x,i})^2 + (y - s_{y,i})^2 \\
        &= x^2 + y^2 -2s_{x,i}x -2s_{y,i}y +s_{x,i}^2 +s_{y,i}^2
    \end{split}\,,
\end{equation}
and the corresponding Jacobians $\mat{H}'_{k,i}$ as
\begin{equation}
    \mat{H}'_{k,i} = 
    \begin{bmatrix}
        2\rv{x}_{k|k-1} - 2s_{x,i} \\
        2\rv{y}_{k|k-1} - 2s_{y,i}
    \end{bmatrix}\,. \label{eqn:modified_jacobian}
\end{equation}
From the above we see that $h'_i(\rvec{x}_{k|k-1})$ and $\mat{H}'_{k,i}$ can be rearranged as linear combinations of the weights $\rv{x}_{k|k-1}^2$, $\rv{y}_{k|k-1}^2$, $\rv{x}_{k|k-1}$ and $\rv{y}_{k|k-1}$, which capture all the required information that is private to the navigator.

>>>>

% Was in problem formulation but all motivations should be in the introduction instead
We motivate these goals with the example of aircraft navigation in the presence of privately-operated range-measuring stations. It is reasonable to assume that the current state of an aircraft may not wish to be disclosed to unknown parties and, similarly, that tower locations may wish to be kept private from unidentified navigating aircraft. The additional safety to passengers provided by accurate aircraft localization, however, may be a goal of all those involved, justifying their cooperation.

@@

\section{Problem Statement} \label{sec:problem_statement}
In this work, we consider the context of privacy-preserving range-sensor navigation, in which we wish for no sensor to learn information about the navigator or other sensors beyond their local measurements, and the navigator to learn no information about individual sensors beyond its location estimate. The problem is two-fold, in that we require explicit cryptographic requirements with a suitable encryption scheme meeting them, as well as an estimation scheme that can use the scheme in the context of range-sensor navigation.

@@

\subsection{Estimation problem} \label{subsec:est_problem}
The estimation problem we consider, for which we will reduce communication to the protocol required for the introduced cryptographic problem, is localisation with range-only sensors. In this work, we will focus on the two-dimensional case for simplicity but will derive methods suitable for extension to a three-dimensional equivalent. The state that we wish to estimate captures the navigator position and the change in position ($\Delta$), and is given by
\begin{equation}
    \vec{x} = 
    \begin{bmatrix}
        x & \Delta x & y & \Delta y
    \end{bmatrix}^\top\,. \label{eqn:state_definition}
\end{equation}
The state evolves following a known system model, which at timestep $k$ is given by
\begin{equation}
    \vec{x}_k = \vec{f}(\vec{x}_{k-1})+\vec{w}_k\,, \label{eqn:system_model}
\end{equation}
with system noise covariance $\mathsf{Cov}[\vec{w}_k]=\mat{Q}_k$. The measurement model is dependent on sensor $i$'s location,
\begin{equation}
    \vec{s}_i = 
    \begin{bmatrix}
        s_{x,i} & s_{y,i}
    \end{bmatrix}^\top\,,
\end{equation} 
and is given by
\begin{equation}
    z_{k,i} = h_i(\vec{x}_k)+v_{k,i}\,, \label{eqn:measurement_model}
\end{equation}
with measurement noise variance $\mathsf{Var}[v_{k,i}] = r_i$ and measurement function $h_i$, for sensor $i$, as
\begin{equation}
    \begin{split}
        h_i(\vec{x}) &= \left\lVert
        \begin{bmatrix}
            x & y
        \end{bmatrix}^\top
        - \vec{s}_{i}\right\rVert \\
        &= \sqrt{(x-s_{x,i})^2 + (y-s_{y,i})^2}\,.
    \end{split}
\end{equation}

Our aim is to provide a filter that estimates the navigator's state $\vec{x}_k$, at every timestep $k$, without learning sensor positions $\vec{s}_i$, measurements $z_{k,i}$ and measurement variances $r_i$ beyond information in the corresponding aggregation leakage function. Similarly, sensors should not learn any information about current state estimates or any other sensor information. Leakage will be further discussed in section \ref{subsec:leakage}, but we note that from any sequential state estimates, following known models, some sensor information leakage can be computed by the navigator. In the context of our leakage function, we will show that this corresponds to the global sums of private sensor information, while individual, or subsets of sensors', information remain private. Similarly, corrupted sensors with access to one or more measurements can produce state estimates of their own, leaking information about navigator state estimates, however, the most accurate estimates, requiring all measurements, will always remain private to the navigator.

@@

Recalling the state definition \eqref{eqn:state_definition} and two-dimensional measurement model \eqref{eqn:measurement_model}, we can see that $h_i$ \textit{cannot} be rearranged to a linear combination of elements private to the navigator. In addition, the Jacobian of $h_i$ at $\rvec{x}_{k|k-1}$,
\begin{equation}
    \mat{H}_{k,i} = 
    \begin{bmatrix}
        \frac{\rv{x}_{k|k-1} - s_{x,i}}{\sqrt{(\rv{x}_{k|k-1} - s_{x,i})^2 + (\rv{y}_{k|k-1} - s_{y,i})^2}} \\
        \frac{\rv{y}_{k|k-1} - s_{y,i}}{\sqrt{(\rv{x}_{k|k-1} - s_{x,i})^2 + (\rv{y}_{k|k-1} - s_{y,i})^2}} 
    \end{bmatrix}\,,
\end{equation}
cannot be either. Therefore, we consider the modified measurement functions
\begin{equation}
    h'_i(\vec{x}) = h_i(\vec{x})^2\,. \label{eqn:modified_measurement_func}
\end{equation}
A measurement function in this form allows rearrangement to a linear combination of navigator information, and thus of the corresponding modified measurement vector $\rvec{i}'_{k,i}$ and measurement matrix $\mat{I}'_{k,i}$ as well. The modified measurement functions $h'_i$ can be written as
\begin{equation}
    \begin{split}
        h'_i(\vec{x}) &= \left\lVert
        \begin{bmatrix}
            x & y
        \end{bmatrix}^\top - \vec{s}_i\right\rVert^2 \\
        &= (x - s_{x,i})^2 + (y - s_{y,i})^2 \\
        &= x^2 + y^2 -2s_{x,i}x -2s_{y,i}y +s_{x,i}^2 +s_{y,i}^2
    \end{split}\,,
\end{equation}
and the corresponding Jacobians $\mat{H}'_{k,i}$ as
\begin{equation}
    \mat{H}'_{k,i} = 
    \begin{bmatrix}
        2\rv{x}_{k|k-1} - 2s_{x,i} \\
        2\rv{y}_{k|k-1} - 2s_{y,i}
    \end{bmatrix}\,. \label{eqn:modified_jacobian}
\end{equation}
From the above we see that $h'_i(\rvec{x}_{k|k-1})$ and $\mat{H}'_{k,i}$ can be rearranged as linear combinations of the weights $\rv{x}_{k|k-1}^2$, $\rv{y}_{k|k-1}^2$, $\rv{x}_{k|k-1}$ and $\rv{y}_{k|k-1}$, which capture all the required information that is private to the navigator.
